### 2018-11-24 tex_files/test.tex
final(ish) presentation

\documentclass[12pt]{article}
\usepackage{smartdiagram}
\usesmartdiagramlibrary{additions}
\begin{document}
	\begin{figure}
		\centering
		\smartdiagramset{
			uniform color list=white!60!green for 3 items,
			back arrow disabled=true,
			module minimum width=2cm,
			module minimum height=2cm,
			module x sep=4cm,
			text width=3cm,
			additions={
				additional item offset=20mm,
				additional item width=2cm,
				additional item height=2cm,
				additional item text width=3cm,
				additional item shadow=drop shadow,
				additional item bottom color=white!60!yellow,
				additional item border color=gray,
				additional arrow color=white!60!yellow,
		}}
		\smartdiagramadd[flow diagram:horizontal]{
			Background Reading, Design ADPLL, Test ADPLL
		}{
			below of module1/External Reference,below of module2/Inverter Chain PLL,below of module3/Interlinked PLLs
		}
		\smartdiagramconnect{->}{additional-module1/additional-module2,additional-module2/additional-module3}
		\begin{tikzpicture}[remember picture,overlay]% modified from p. 47 of manual
		\draw[additional item arrow type] (module3) |- ([yshift=-10mm]module1.south) -- (additional-module1);
		\end{tikzpicture}

		\vspace{40mm}\vspace{20mm}
		\smartdiagramset{
			uniform color list=white!60!orange for 6 items,
			back arrow disabled=true,
			module minimum width=2cm,
			module minimum height=2cm,
			module x sep=4cm,
			text width=3cm,
			additions={
				additional item offset=20mm,
				additional item width=2cm,
				additional item height=2cm,
				additional item text width=3cm,
				additional item shadow=drop shadow,
				additional item bottom color=white!60!black,
				additional item border color=gray,
				additional arrow color=orange,
		}}
		\begin{tikzpicture}[remember picture,overlay]% modified from p. 47 of manual
		\draw[additional item arrow type] (additional-module3) |- ([yshift=-10mm]additional-module1.south) -- (additional-module1);
		\end{tikzpicture}
		\smartdiagramadd[flow diagram:horizontal]{
			Acquire Better FPGA, Build ADPLL Network,  Test/Analyse Network
		}{
		below of module1/External Reference,below of module2/Inverter Chain PLL,below of module3/Interlinked PLLs
	}\par
	\end{figure}
\end{document}