### 2019-04-15 thesis/Conclusion.tex
48.5

\section{Conclusions}
\section{Impact of Research}
\section{Suggested Future Work}