### 2019-04-15 thesis/TestingAnalysis.tex

\section{2x2 Network Performance Comparison}
\section{3x3 Network Performance Comparison}