### 2019-04-21 thesis/CodeAppendix.tex

A selection of files containing the core Verilog modules are included in this appendix. All files used can be located on github at  \href{https://github.com/ConchuOD/fpga_adpll}{\textbf{https://github.com/ConchuOD/fpga\_adpll}}.
\section{Phase Detector}\label{lst:PhaseDetector}
\lstinputlisting[style={verilog-style}]{../design_sources/PhaseDetector.v}
\lstinputlisting[style={verilog-style}]{../design_sources/UpDownCounter.v}

\section{Phase Detector with \acs{TDL}}\label{lst:phase_detector_dl}
\lstinputlisting[style={verilog-style}]{../design_sources/PhaseDetectorDL.v}

\section{Error Combiner}\label{lst:error_combiner}
\lstinputlisting[style={verilog-style}]{../design_sources/ErrorCombiner.v}

\section{Loop Filter}\label{lst:loop_filter}
\lstinputlisting[style={verilog-style}]{../design_sources/LoopFilter.v}

\section{Ring Oscillator}\label{lst:ring_osc}
\lstinputlisting[style={verilog-style}]{../design_sources/RingOsc.v}

\section{\acs{FPGA} Clocked Oscillator}\label{lst:phase_accum}
\lstinputlisting[style={verilog-style}]{../design_sources/PhaseAccum.v}

\section{Network \acs{ADPLL} (\acs{RO})}\label{lst:network_ring_adpll}
\lstinputlisting[style={verilog-style}]{../design_sources/NetworkRingADPLL.v}

\section{Network \acs{ADPLL} (\ac{FPGA} Clocked)}\label{lst:network_adpll}
\lstinputlisting[style={verilog-style}]{../design_sources/NetworkRingADPLL.v}